---
title: "Network of Contributors"
output:
  html_document:
    self_contained: TRUE
    keep_md: FALSE
---

The interactive network visualization below shows the connections between FORRT contributors based on their collaborative work across different projects. Each node represents a contributor, and the connections (edges) represent shared project participation.

<style>
.network-container {
  display: flex;
  justify-content: center;
  align-items: center;
  width: 100%;
  height: 100vh;
  margin: 0 auto;
}

.vis-network {
  margin: 0 auto;
  display: block;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE, 
  fig.path = ""
)
```

```{r load-libraries}
library(visNetwork)
library(dplyr)
library(tidyr)
library(igraph)
library(googlesheets4)
library(stringr)
```

```{r data-processing}
# Google Sheet URLs
google_sheet_url_automation <- "https://docs.google.com/spreadsheets/d/1MUD54FQUhfcBKrvr5gCYoh2wgbJ6Lf7oAJRAqsQ-Nag/edit"
google_sheet_url_leads <- "https://docs.google.com/spreadsheets/d/1roy-sZTxyXENA5c5IIV7IIemYvzbzs7zojUN2yIpi58/edit?gid=0#gid=0"

# Get sheet names
gs4_deauth()
sheets <- sheet_names(google_sheet_url_automation)

# Exclude the first four Automation Source sheets
relevant_sheets <- sheets[-c(1,2,3,4)]

# Exclude 'full' sheets
relevant_sheets <- relevant_sheets[!grepl(" - full$", relevant_sheets)]

# Read all relevant sheets into a list and remove rows without last names
internal_links <- read_sheet(google_sheet_url_automation, sheet = "INTERNAL LINKS") %>% 
  select(`Project Name`) # Extract project names

data_list <- lapply(relevant_sheets, function(sheet) {
  # Read the content of the current sheet
  df <- read_sheet(google_sheet_url_automation, sheet = sheet)
  # Remove rows where the `Surname` is missing or empty
  df <- df %>% filter(!is.na(`Surname`) & `Surname` != "")
  # Match the `Project Name` using the sheet name (assuming sheet name is the identifier)
  project_name <- internal_links$`Project Name`[internal_links$`Project Name` == sheet]
  # If a match is found, assign the `Project Name`, else set to NA
  if (length(project_name) > 0) {
    df <- df %>% mutate(`Project Name` = project_name)
  } else {
    df <- df %>% mutate(`Project Name` = NA)
  }
  df
})

# Combine all sheets into one dataframe
dt <- bind_rows(data_list, .id = "Source_Sheet")

# Move project name column to first position 
dt <- dt %>%
  select(`Project Name`, everything())

# Remove PM column as it is unnecessary and might cause problems with binding
dt <- dt %>%
  select(-`Project Managers`)
```

```{r leads-data}
# Read Leads Tenzing data
leads_sheets <- sheet_names(google_sheet_url_leads)
relevant_leads_sheets <- leads_sheets[-c(2)]
leads_df <- read_sheet(google_sheet_url_leads)

# Rename columns and select relevant ones
leads_df <- leads_df %>%
  rename(
    `Project Name` = `FORRT project(s)`,
    `ORCID iD` = `ORCID`
  ) %>%
  select(`First name`, `Middle name`, `Surname`, `Project Name`, `Role`, `ORCID iD`)

# Create a column to mark the presence of a role and group by individual
leads_df <- leads_df %>%
  mutate(has_role = TRUE) %>%
  distinct(`First name`, `Middle name`, `Surname`, `Project Name`, `ORCID iD`, `Role`, .keep_all = TRUE)

# Pivot the data to create a column for each leadership role type, removing director
leads_df <- leads_df %>%
  pivot_wider(names_from = `Role`, values_from = `has_role`, values_fill = list(has_role = FALSE)) %>%
  select(-Director, -`Operations Coordinator`)

# Combine leads tenzing rows with automation source
dt <- bind_rows(dt, leads_df)
```

```{r data-cleaning}
# Trim names and ORCIDs to ensure inconsistent Tenzing entries are not counted separately 
dt <- dt %>%
  mutate(
    `First name` = str_trim(str_replace_all(`First name`, "\\*", "")),
    `Middle name` = str_trim(str_replace_all(`Middle name`, "\\*", "")) %>% str_sub(1, 1),
    `Surname` = str_trim(str_replace_all(`Surname`, "\\*", "")),
    `ORCID iD` = str_trim(str_remove(`ORCID iD`, "https://orcid.org/"))
  )

# Ensure Conceptualization is logical
dt$Conceptualization <- as.logical(dt$Conceptualization)
```

```{r network-data-preparation}
# Process data for network
dt_long <- dt %>%
  pivot_longer(cols = c("Conceptualization", "Data curation", "Formal analysis", "Funding acquisition", "Investigation", "Methodology", "Project administration", "Resources", "Software", "Supervision", "Validation", "Visualization", "Writing - original draft", "Writing - review & editing"), 
              names_to = "Role", 
              values_to = "has_role") %>%
  filter(has_role == TRUE) %>%
  rowwise() %>% 
  mutate(Contributor = paste(na.omit(c(`Surname`, `First name`)), collapse = " ")) %>% 
  mutate(Lead = if_else(Role %in% c("lead", "co-lead"), Role, "other"))

# Get co-occurences
Contributor_Project <- xtabs(~ `Project Name` + Contributor, 
                            data=dt_long %>% 
                              select(`Project Name`, Contributor) %>% 
                              unique(), 
                            sparse = TRUE)

Contributor_occur <- crossprod(Contributor_Project, Contributor_Project)
```

```{r create-nodes}
# Create nodes
Contributor_nodes_n <- dt_long %>%
  group_by(Contributor) %>% 
  count(Contributor)

colnames(Contributor_nodes_n) <- c("id", "value")

ContributorGroups_nodes <- dt_long %>% 
  select(Contributor, `Project Name`) %>% 
  unique() %>% 
  group_by(Contributor) %>% 
  mutate(Projects = paste0(`Project Name`, collapse = ",")) %>% 
  select(Contributor, Projects) %>% 
  unique()

colnames(ContributorGroups_nodes) <- c("id", "projects")

Contributor_nodes <- left_join(Contributor_nodes_n, ContributorGroups_nodes)
Contributor_nodes$title <- paste0(Contributor_nodes$id,"<br>", gsub(pattern = ",", replacement = "<br>", x = Contributor_nodes$projects))
```

```{r create-edges}
# Create edges
Contributor_edges <- which(upper.tri(Contributor_occur), arr.ind = TRUE) %>%
  as.data.frame() %>%
  mutate(
    from = rownames((Contributor_occur))[row],
    to   = colnames((Contributor_occur))[col],
    width_n = (Contributor_occur)[cbind(row, col)]
  ) %>%
  filter(width_n > 0) %>%
  mutate(width = (width_n/10)^(3)) %>% 
  select(from, to, width) 

# Add one hidden edge that will move the Chinese translator team closer to the center
additional_hidden_edge <- data.frame(from = c("Fang Cathy", "Chen Liangjie", "Jin Shuxian", "Yang Jinbiao", "Liu Ruoting", "Wang Xinyu", "Xu Yu", "Ji Xuejun", "Wang Zixi"), 
                                     to = c("Azevedo Flavio"), 
                                     width = 0.0,
                                     hidden = TRUE)

Contributor_edges_plot <- rbind(Contributor_edges %>% 
                                 mutate(hidden = F), additional_hidden_edge)
```

```{r network-visualization, echo=FALSE}
# Create the network visualization
visNetwork(Contributor_nodes, Contributor_edges_plot, width = "100%", height = "800px") %>%
  visLayout(randomSeed = 1001) %>%
  visIgraphLayout(layout = "layout_with_fr", randomSeed = 1001) %>%
  visPhysics(
    enabled = TRUE,
    stabilization = list(
      enabled = TRUE,
      iterations = 2000,
      fit = FALSE
    ),
    solver = "forceAtlas2Based",
    forceAtlas2Based = list(
      gravitationalConstant = -80,   # Light repulsion to preserve central clusters
      centralGravity = 0.4,          # Moderate central pull for distant nodes
      springLength = 150,            # Medium springs - shorter for distant nodes
      springConstant = 0.003,        # Moderate spring strength
      damping = 0.4,                 # Balanced damping
      avoidOverlap = 0.2             # Light overlap avoidance to preserve density
    )
  ) %>%
  visEdges(color = list(color = "#87CEEB", opacity = 0.3)) %>%
  visNodes(
    font = list(size = 4, strokeWidth = 1, strokeColor = "white"),
    size = 7,
    borderWidth = 0,
    color = list(background = "#000000", highlight = list(background = "#A52828", border = "darkred")),
    scaling = list(min = 3, max = 15)
  ) %>%
  visInteraction(
    dragNodes = TRUE,
    dragView = TRUE,
    hover = TRUE,
    hoverConnectedEdges = TRUE,
    selectable = TRUE,
    multiselect = TRUE,
    navigationButtons = TRUE
  ) %>%
  visOptions(
    highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE, algorithm = "hierarchical"),
    nodesIdSelection = list(enabled = TRUE, style = "width: 200px;"),
    selectedBy = list(variable = "projects", multiple = TRUE, style = "width: 200px; background: #f8f8f8;")
  ) %>%
  htmlwidgets::onRender("
    function(el, x) {
      var network = this;
      
      // Set initial zoom level immediately
      network.on('initRedraw', function() {
        network.moveTo({
          scale: 0.012,
          animation: false
        });
      });
      
      // Maintain zoom after stabilization
      network.on('stabilized', function() {
        network.moveTo({
          scale: 0.012,
          animation: false
        });
      });
      
      network.on('click', function(params) {
        if (params.nodes.length === 0) {
          network.body.data.nodes.update(
            network.body.data.nodes.getIds().map(function(id) {
              return {id: id, color: {background: '#353839'}};
            })
          );
        }
      });
      
      network.on('selectNode', function(params) {
        network.body.data.nodes.update(
          network.body.data.nodes.getIds().map(function(id) {
            return {id: id, color: {background: '#353839'}};
          })
        );
        network.body.data.nodes.update({
          id: params.nodes[0],
          color: {background: '#A52828'}
        });
      });
    }
  ")
```
